### 2025-03-03 README.Rmd
Fixes for several issues; mostly copy edits (#24)
* Copy edits fix #15

* Revise structure and add new fixes

* Copy eidts to fix #17

* Copy edits; closes #18

* Copy edits; closes #19

* Adding minimum versions for dependencies

* Remove minimum on base R packages; closes #16
update install instructions, title, desc

# hatchR <a href="https://bmait101.github.io/hatchR/"><img src="man/figures/logo.png" align="right" height="139" alt="hatchR website" /></a>

@@

**hatchR** is an R package for predicting fish developmental phenology using statistical models. It is intended to bridge the analytic gap of taking statistical models developed in hatchery settings (*e.g.*, Beacham and Murray 1990) and applying them to real world temperature data sets using the effective value framework developed by Sparks et al. (2019).

@@

You can install the development version of **hatchR** from [GitHub](https://github.com/) with:

```{r}
# install.packages("pak")
pak::pak("bmait101/hatchR")
```

@@

>>>>

# hatchR: Predict Fish Hatch and Emergence Timing <a href="https://bmait101.github.io/hatchR/"><img src="man/figures/logo.png" align="right" height="139" alt="hatchR website" /></a>

@@

**hatchR** is an R package for predicting hatch and emergence timing for a wide range of wild fishes using the effective value framework (Sparks et al., (2019) <DOI:10.1139/cjfas-2017-0468>). 'hatchR' offers users access to established phenological models and the flexibility to incorporate custom parameterizations using external datasets. The package includes functions for model selection, phenology prediction, and visualization. **hatchR** is designed to be user-friendly and accessible to a wide range of users, including researchers, managers, and educators.

@@

Install and load the most recent approved version from CRAN by running:

```{r}
# install the most recent approved version from CRAN
install.packages("hatchR")
```

Install the development version of **hatchR** from [GitHub](https://github.com/) by running:

```{r}
# install the most recent version from GitHub
pak::pak("bmait101/hatchR")
```

## Interactive Shiny App

You can also use the **hatchR** Shiny app to interactively predict fish developmental phenology. To run the app, visit the [hatchR Shiny app](https://elifelts.shinyapps.io/hatchR_demo/).

@@

**hatchR** requires three primary data inputs:

1.  Water temperature data
2.  Species-specific model parameters
3.  Spawning date or date ranges
